\vspace{-1cm}\normalsize{\textbf{Fairly Random: Impact of Winning the Toss on the Probability of Winning}}

### 2015-12-06 write_up/cricket.tex
nip/tuck

We assume that the outcome of a toss is random. Conditional on the outcome of a toss being random, the effect of winning the toss can be attributed to the toss itself. Before we exploit the design, we shed some light on the validity of the assumption. In particular, we assess whether the coin toss is somehow rigged, with home sides winning tosses more often. For this task, we limited ourselves to only international matches as establishing which of the teams is the home team in local matches is somewhat arduous. Of the 4,749 international matches, the home team won the toss in 2,433 matches, or about 51.2\%. The probability that we will get as large a proportion by chance after tossing close to 5,000 coins is about 9\%, low but not eyebrow-raisingly so.

In cricket matches, three results are possible: a draw, a win, and a loss. We look at draws, a common outcome in matches in which overs aren't limited, later on. For now, we focus on winning and losing. We elide over differential number of draws across formats, estimating the difference in probability of winning and losing.

>>>>

We assume that the outcome of a toss is random. Conditional on the outcome of a toss being random, the effect of winning the toss can be attributed to the toss itself. Before we exploit the design, we shed some light on the validity of the assumption. In particular, we assess whether the coin toss is somehow rigged, with the home side enjoying the rub of the green more often. For this analysis, we only get to exploit international matches as establishing which of the teams is the home team in local matches is somewhat arduous. Of the 4,749 international matches, the home team won the toss in 2,433 matches, or about 51.2\%. The probability that we will get as large a proportion by chance after tossing close to 5,000 coins is about 9\%, low but not eyebrow-raisingly so. Thus, for now, we choose to trust that the tosses are random.

In cricket matches, three results are possible: a draw, a win, and a loss. For now, we focus on winning and losing. We look at draws, a common outcome in matches in which overs aren't limited, later on. We elide over differential number of draws across formats, estimating the difference in probability of winning and losing.
